t---
title: 'Trade Prevalidation: LM residuals'
author: "Aleksandr Matrunich"
date: "04/27/2015"
output: html_document
---

For testing purposes we take inflows of two diverse reporters (like US and UZ) and test on them detection of outliers with regression linear model.

```{r}
faosws::GetTestEnvironment(
  baseUrl = "https://hqlprswsas1.hq.un.fao.org:8181/sws",
  token = "11ac4873-4747-43cd-b942-711d0ebfe844"
  )
```

Reporters
```{r}
getCountryCode("USA\\(1981-")
getCountryCode("Uzb")
```

Elements
```{r}
selectElems(direction == "in" &
              unitgroup %in% c("weight", "cost", "price", "volume"))
```


Items
```{r}
getAllItems() %>%
  filter(str_detect(code, "^10") &
           str_length(code) == 6 &
           str_detect(description, "Durum wheat")) %>%
  select(code) %>%
  unlist()
```

Data
```{r}
data <- getComtradeData(reporter = c(getCountryCode("USA\\(1981-")$code,
                                     getCountryCode("Germany\\(1")$code),
                        partner = getAllPartnersRaw() %>%
                          normalizeAreas() %>%
                          filter(type == "country") %>%
                          select(code)  %>% unlist(),
                        year = 1990:2014, 
                        item = getAllItems() %>%
                          filter(str_detect(code, "^10") &
                                   str_length(code) == 6 &
                                   str_detect(description, "Durum wheat")) %>%
                          select(code) %>%
                          unlist(),
                          element = selectElems(direction == "in" &
                                                  unitgroup %in% c("weight", 
                                                                   "cost", 
                                                                   "price", 
                                                                   "volume"))) %>%
  humanizeComtrade()

```

Are there only kg and $US?
No retrade?

```{r}
all(unique(data$unit) %in% c("kg", "US$"))
all(unique(data$back) %in% FALSE)
```

## Building models

### Tests

```{r}
data %>%
  group_by(reporter, partner, year, dir, back, hs) %>%
  mutate(n = n()) %>%
  top_n(10, n)
```

### How to calculate price?
```{r}

## With reshape2
data %>%
  select(-unit) %>% # We drop unit but we should to check unit consistensy before
  reshape2::dcast(... ~ group) %>%
  mutate(price = cost / weight)

## With tidyr
data %>%
  select(-unit) %>%
  tidyr::spread(group, value) %>%
  mutate(price = cost / weight)


```

### Build models
```{r}
outers <- data %>%
  select(-unit) %>% # We drop unit but we should to check unit consistensy before
  reshape2::dcast(... ~ group) %>%
  mutate(price = cost / weight) %>% 
  group_by(reporter, partner, dir, back, item, hs) %>%
  do(broom::augment(lm(price ~ year, data = .))) %>%
  ungroup() %>%
  arrange(desc(abs(.std.resid))) %>%
  select(reporter, partner, dir, back, item, hs, .std.resid) %>%
  top_n(20, .std.resid) %>%
  select(-.std.resid) %>%
  distinct()
  
```

### Let's draw!

Data for plotting

```{r}
outers %>%
  left_join(data)  %>% 
  select(-unit) %>% # We drop unit but we should to check unit consistensy before
  reshape2::dcast(... ~ group) %>%
  mutate(price = cost / weight) %>% 
  ggplot(aes(as.Date(paste0(year, ".12.31"), "%Y.%m.%d"), price, group = 1)) +
  geom_point(aes(size = weight / 1000)) +
  geom_smooth(method = "lm", se = F) + 
  facet_wrap(~ partner, scales = "free") +
  scale_size_continuous("Weight, t", trans = "log10", labels = scales::comma) +
  scale_y_continuous("Price, US$/kg", labels = scales::dollar) +
  scale_x_date("")
```

We need to change year to date class to discover gaps in data: as.Date("1996", "%Y")


Problems with Germany and Italy

```{r}
getComtradeData(reporter = getCountryCode("USA\\(1981-")$code,
                partner = getCountryCode("Germany\\(1")$code,
                year = 1990:2014, 
                item = getAllItems() %>%
                  filter(str_detect(code, "^10") &
                           str_length(code) == 6 &
                           str_detect(description, "Durum wheat")) %>%
                  select(code) %>%
                  unlist(),
                element = selectElems(direction == "in" &
                                        unitgroup %in% c("weight", 
                                                         "cost", 
                                                         "price", 
                                                         "volume"))) %>%
  humanizeComtrade()
```


getTrade("USA", "Germany", 1990:2014, items("Durum wheat"))

```{r}

data <- getComtradeData(reporter = c(getCountryCode("USA\\(1981-")$code,
                                     getCountryCode("Germany\\(1")$code),
                        partner = getAllPartnersRaw() %>%
                          normalizeAreas() %>%
                          filter_(.dots = list(~type == "country")) %>%
                          select_(~code)  %>% unlist(),
                        year = 1990:2014, 
                        item = getAllItems() %>%
                          filter_(.dots = list(~stringr::str_detect(code, "^10") &
                                                 stringr::str_length(code) == 6)) %>%
                          select_(~code) %>% unlist(),
                        element = selectElems(direction == "in" &
                                                unitgroup %in% c("weight", 
                                                                 "cost", 
                                                                 "price", 
                                                                 "volume"))) %>%
  humanizeComtrade()

if(any(!is.element(unique(data$unit), c("kg", "US$")))) stop("Other than kg and US$ units present!")




outers <- data %>%
  select_(.dots = list(~-unit)) %>% # We drop unit but we should to check unit consistensy before
  reshape2::dcast(... ~ group) %>%
  mutate_(.dots = list(price = ~cost / weight)) %>% 
  group_by_(~reporter, ~partner, ~dir, ~back, ~item, ~hs) %>%
  do(broom::augment(lm(price ~ year, data = .))) %>% # NSE here!
  ungroup() %>%
  arrange_(~desc(abs(.std.resid))) %>%
  select_(~reporter, ~partner, ~dir, ~back, ~item, ~hs, ~price, ~year, ~.std.resid) %>%
  top_n(20, .std.resid) %>%
  select_(~-.std.resid) %>%
  distinct()


```

Problem with humanizeComtrade
Warning message:
In `[.data.table`(`_dt`, , `:=`(item, getHSDesc(measuredItemHS)$description),  :
  Supplied 33 items to be assigned to 8304 items of column 'item' (recycled leaving remainder of 21 items).
  
Done.
  
```{r}
getHSDesc(data$measuredItemHS)
```

```{r}
broom_augment <- function(model) {
  cbind(model.frame(model), .std.resid = rstandard(model))
  }
```



